---
title: "Interactions analyses with CellChat on LPS condition"
author: "Matei Teleman"
date: "11/3/2021"
output: html_document
---

## Import libraries
```{r}
library(CellChat)
library(ggplot2)                  
library(patchwork)
library(ggalluvial)
library(igraph)
library(dplyr)
library(Seurat)
```


## Preparing for the data for the Nichenet analyses
### Importation of .RDS file of meningeal dataset after QC and demultiplexing and annotation of the clusters
```{r}
so <- readRDS("/mnt/DOSI/PLATEFORMES/BIOINFORMATIQUE/03_WORKSPACE/teleman/ciml-rrlab/Project/TotalMeningeLPS/202010028_R333_RR/01_Input/Seurat objects/TotalMeningeLPS_202010028_R333_RR_sc10x.rna.seurat_50PCA.RDS")
DimPlot(so,label = T,reduction = "umap1")
new.cluster.ids <- c("Stromal", "Stromal", "Stromal","Stromal","Stromal","Stromal","B cell","Macrophages","Endothelium"
                     , "Neutrophils Monocytes", "Mast cell","Activated Macrophages", "NK","Stromal",
                     "ILC2","DC/pDC", "Osteoblasts", "Endothelium","Schwann","B cell", "MigDC","NK","Neuron","Unknown",
                     "Stromal","Neutrophils Monocytes","Smooth muscle","Lymphatics","B cell")
names(new.cluster.ids) <- levels(so)
so <- RenameIdents(so, new.cluster.ids)
DimPlot(so,label = T,reduction = "umap1")
```

### adding MM MHCII+ & MM MHCII- in annotations
#### Clustering the macrophages clusters after subsetting them
```{r}
sbmacs <- subset(so, idents = c("Macrophages","Activated Macrophages"))
sbmacs <- SetIdent(sbmacs, value = sbmacs[["MULTI_Sample"]])
sbmacs <- FindNeighbors(sbmacs, dims = 1:50)
sbmacs <- FindClusters(sbmacs, resolution = 0.3,method = 1)
sbmacs <- RenameIdents(sbmacs,"0" = "MM MHCII-","1" = "MM MHCII+","2" = "MM MHCII+","3" = "MM MHCII-")
DimPlot(sbmacs,reduction = "umap1")
```

#### Removing the macrophages clusters from the original seurat object of the meninges
```{r}
so.sans.macs <- subset(so, idents = unique(c("Stromal", "Stromal", "Stromal","Stromal","Stromal","Stromal","B cell","Endothelium"
                     , "Neutrophils Monocytes", "Mast cell", "NK","Stromal",
                     "ILC2","DC/pDC", "Osteoblasts", "Endothelium","Schwann","B cell", "MigDC","NK","Neuron","Unknown",
                     "Stromal","Neutrophils Monocytes","Smooth muscle","Lymphatics","B cell")))
DimPlot(so.sans.macs)
```

#### Merging the two resulting seurat objects
```{r}
so.v2 <- merge(so.sans.macs,sbmacs,merge.data = T,merge.dr = "umap1")
DimPlot(so.v2)
```

#### Subsetting the LPS-only cells
```{r}
so.v2.lps <- subset(so.v2, subset = MULTI_Sample == "LPS")
DimPlot(so.v2.lps,reduction = "umap1")
```

### Preparing CellChat datasets
```{r}
data.input <- GetAssayData(so.v2.lps, assay = "RNA", slot = "data") # normalized data matrix
labels <- Idents(so.v2.lps)
meta <- data.frame(group = labels, row.names = names(labels)) # create a dataframe of the cell labels
cellchat <- createCellChat(object = data.input, meta = meta, group.by = "group")
cellchat <- addMeta(cellchat, meta = meta, meta.name = "labels")
cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels
```

```{r}
CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)
```


```{r}
# Show the structure of the database
dplyr::glimpse(CellChatDB$interaction)

# use a subset of CellChatDB for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling") # use Secreted Signaling
# use all CellChatDB for cell-cell communication analysis
CellChatDB.use <- CellChatDB # simply use the default CellChatDB

# set the used database in the object
cellchat@DB <- CellChatDB.use
```

## Performing CellChat analyses
```{r}
cellchat <- subsetData(cellchat) # subset the expression data of signaling genes for saving computation cost
#future::plan("multiprocess", workers =12) # do parallel

cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
cellchat <- projectData(cellchat, PPI.mouse)
```

```{r}
cellchat <- computeCommunProb(cellchat, raw.use = TRUE)
# Filter out the cell-cell communication if there are only few number of cells in certain cell groups
cellchat <- filterCommunication(cellchat, min.cells = 10)
```

```{r}
cellchat <- computeCommunProbPathway(cellchat)
cellchat <- aggregateNet(cellchat)
groupSize <- as.numeric(table(cellchat@idents))
```

## Visualization of the results
### Global visual
```{r}
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

### Visualization for each meningeal population
```{r}
mat <- cellchat@net$weight
dim(mat)
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

### Visualization for macrophages interactions
In clust2 vector, change 7 by 8 for the other population. 7 stands for MM MHCII- and 8 for MM MCHII+.
```{r}
filtered_matrix <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
clusts2 <- c(8)#,7)
for (c in clusts2){
  filtered_matrix[,c] <- mat[, c]
  filtered_matrix[c,] <- mat[c,]
}

netVisual_circle(filtered_matrix,vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat))
```

```{r}
netVisual_bubble(cellchat, sources.use = c(7,8),targets.use = c(12), remove.isolate = TRUE,font.size =10)
netVisual_bubble(cellchat, sources.use = c(12),targets.use = c(7,8), remove.isolate = TRUE,font.size =10)
```