---
title: "Clustering of the Macrophages after sub-setting"
author: "Matei Teleman"
date: "11/3/2021"
output: html_document
---


## Import libraries
```{r}
library(CellChat)
library(ggplot2)
library(ggalluvial)
library(igraph)
library(dplyr)
library(Seurat)
library(tidyverse)
library("writexl")
library(clusterProfiler)
library(enrichplot)
organism = "org.Mm.eg.db"
BiocManager::install(organism, character.only = TRUE)
options(connectionObserver = NULL)
library(organism, character.only = TRUE)
library(BiocParallel)
library(pathview)
library(EnhancedVolcano)
```


## Create function for GSEA with GO database
```{r}

GO_analysis <- function(data,pn,eps,fs){
  # we want the log2 fold change 
  original_gene_list <- data$avg_log2FC
  
  # name the vector
  names(original_gene_list) <- data$gene
  
  # omit any NA values 
  gene_list<-na.omit(original_gene_list)
  gene_list <- gene_list[!is.na(gene_list)]
  # sort the list in decreasing order (required for clusterProfiler)
  gene_list = sort(gene_list, decreasing = TRUE)
  #print(gene_list)
  gse <- gseGO(geneList=gene_list, 
             ont ="ALL", 
             keyType = "SYMBOL", 
             #nPerm = 10000, 
             minGSSize = 3, 
             maxGSSize = length(gene_list)-1, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = org.Mm.eg.db,
             scoreType = pn,
             pAdjustMethod = "none",eps = eps)
  require(DOSE)
  return(gse)
  }
```

## Importation of .RDS file of meningeal dataset after QC and demultiplexing
```{r}
s_o <- readRDS("/mnt/DOSI/PLATEFORMES/BIOINFORMATIQUE/03_WORKSPACE/teleman/ciml-rrlab/Project/TotalMeningeLPS/202010028_R333_RR/01_Input/Seurat objects/TotalMeningeLPS_202010028_R333_RR_sc10x.rna.seurat_50PCA.RDS")
DimPlot(s_o,reduction = "umap1", label = T)
s_o <- FindClusters(s_o, resolution = 1,method = 1)
DimPlot(s_o,reduction = "umap1", label = T)
# Plot of C1qa (marker gene of macrophages in meninges)
FeaturePlot(s_o,reduction = "umap1",features = "C1qa" )
# Plot of CD206/Mrc1
FeaturePlot(s_o,reduction = "umap1",features = "Mrc1" )
```


## Sub-setting of both macrophages cluster
```{r}
sbmacs <- subset(s_o, idents = c("7","11"))
DimPlot(sbmacs, reduction = "umap1")
# Plot of meninges cell population depending of their conditions : PBS or LPS
DimPlot(sbmacs, reduction = "umap1",group.by = "MULTI_Sample")
# Plot of C1qa (marker gene of macrophages in meninges)
FeaturePlot(sbmacs,"C1qa",reduction = "umap1")
```

## Clustering of the macrophages subset
```{r}
# Setting study condition (LPS or PBS) as active identity
sbmacs <- SetIdent(sbmacs, value = sbmacs[["MULTI_Sample"]])
DimPlot(sbmacs,reduction = "umap1")
# Clustering
sbmacs <- FindNeighbors(sbmacs, dims = 1:50)
sbmacs <- FindClusters(sbmacs, resolution = 0.3,method = 1)
# Ploting new clustering
DimPlot(sbmacs,reduction = "umap1")
DimPlot(sbmacs,reduction = "umap1",group.by = "MULTI_Sample")
# Ploting a gene marker of MHC-II
VlnPlot(sbmacs,"H2-Ab1")
FeaturePlot(sbmacs,reduction = "umap1",features = "H2-Ab1")
VlnPlot(sbmacs,"H2-Ab1",split.by = "MULTI_Sample")
```


## Computing DEGs for both macrophages populations
With the obtained clustering, we assign clusters :
  - 0 : MM CD206+ MHCII- PBS
  - 1 : MM CD206+ MHCII+ PBS
  - 2 : MM CD206+ MHCII+ LPS
  - 3 : MM CD206+ MHCII- LPS
  
### Computing MM CD206+ MHCII+ DEGs between LPS and PBS conditions
```{r}
dpos.degs <- FindMarkers(sbmacs,ident.1 = "2",ident.2 = "1",min.pct = 0.25) %>% rownames_to_column("gene") %>% filter(p_val_adj <= 0.05)
head(dpos.degs)
# Saving locally output results
write_xlsx(dpos.degs,path = "path/MACS_2POS_DEGs_LPSvsPBS_clustering.xlsx")
```

### Computing MM CD206+ MHCII- DEGs between LPS and PBS conditions
```{r}
dneg.degs <- FindMarkers(sbmacs,ident.1 = "3",ident.2 = "0",min.pct = 0.25) %>% rownames_to_column("gene") %>% filter(p_val_adj <= 0.05)
head(dneg.degs)
# Saving locally output results
write_xlsx(dneg.degs,path = "path/MACS_2NEG_DEGs_LPSvsPBS_clustering.xlsx")
```

## Analyzing DEGs output
### GSEA Analysis of MM CD206+ MHCII+ DEGs on GO database
```{r}
mdpos.cl.pathways <- GO_analysis(dpos.degs,NULL,1e-10,15)
# Ploting top 10 pathways
dotplot(mdpos.cl.pathways, showCategory=10, split=".sign") + facet_grid(.~.sign)
```

### GSEA Analysis of MM CD206+ MHCII- DEGs on GO database
```{r}
mdneg.cl.pathways <-GO_analysis(dneg.degs,NULL,1e-10,15)
# Ploting top 10 pathways
dotplot(mdneg.cl.pathways, showCategory=10, split=".sign") + facet_grid(.~.sign)
```

### Volcano plot of MM CD206+ MHCII+ DEGs on GO database
```{r}
EnhancedVolcano(dpos.degs,lab = dpos.degs$gene,x = 'avg_log2FC',y = 'p_val')
```

### Volcano plot of MM CD206+ MHCII- DEGs on GO database
```{r}
EnhancedVolcano(dneg.degs,lab = dneg.degs$gene,x = 'avg_log2FC',y = 'p_val')
```

## Computing DEGs for macrophages CD206+
```{r}
macs.degs <- FindMarkers(sbmacs,ident.1 = c("3","2"),ident.2 = c("1","0"),min.pct = 0.25)%>% rownames_to_column("gene") %>% filter(p_val_adj <= 0.05)
head(macs.degs)
write_xlsx(macs.degs,path = "path/TOTAL_MACS_CD206_DEGs_LPSvsPBS_clustering.xlsx")
```

### Plot Volcano plot of MM CD206+ DEGs
```{r}
EnhancedVolcano(macs.degs,lab = macs.degs$gene,x = 'avg_log2FC',y = 'p_val')
```

### GSEA Analysis of MM CD206+ DEGs on GO database
```{r}
GO_analysis(macs.degs,NULL,1e-10,12)
```

## Venn diagram between MM CD206+ MHCII+ & MM CD206+ MHCII- DEGs
```{r}
ggvenn(list(`Macs CD206+ MHC-II+` = dpos.degs$gene,`Macs CD206+ MHC-II-` = dneg.degs$gene),set_name_size = 12,show_percentage = F,text_size = 15)
```

## Venn diagram between MM CD206+ MHCII+ & MM CD206+ MHCII- DEGs (filtered with the corresponding thresholds of Volcano plots)
```{r}
filt_degs2pos <- subset(dpos.degs,subset = abs(avg_log2FC) > 1 & p_val_adj < 1e-5)
filt_degs2neg <- subset(dneg.degs,subset = abs(avg_log2FC) > 1 & p_val_adj < 1e-5)
```

```{r}
ggvenn(data = list(`Macrophages MHCII+` = filt_degs2pos$gene,`Macrophages MHCII-` = filt_degs2neg$gene),show_percentage = F,set_name_size = 4,text_size = 9,)
```