---
title: "Monocle on Meningeal Macrophages"
author: "Matei Teleman"
date: "21/07/2022"
output: html_document
---

## Import libraries
```{r}
library(monocle3)
library(Seurat)
library(SeuratData)
#library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(magrittr)
library(tidyverse)
```


## Preparing for the data for the Monocle analyses
### Importation of .RDS file of meningeal dataset after QC and demultiplexing and subsetting macrophage clusters
```{r}
s_o <- readRDS("/mnt/DOSI/PLATEFORMES/BIOINFORMATIQUE/03_WORKSPACE/teleman/ciml-rrlab/Project/TotalMeningeLPS/202010028_R333_RR/01_Input/Seurat objects/TotalMeningeLPS_202010028_R333_RR_sc10x.rna.seurat_50PCA.RDS")
DimPlot(s_o,reduction = "umap1", label = T)
#s_o <- FindNeighbors(s_o, dims = 1:50)
s_o <- FindClusters(s_o, resolution = 1,method = 1)
DimPlot(s_o,reduction = "umap1", label = T)
FeaturePlot(s_o,reduction = "umap1",features = "C1qa" )
```


```{r}
sbmacs <- subset(s_o, idents = c("7","11"))
DimPlot(sbmacs, reduction = "umap1")
DimPlot(sbmacs, reduction = "umap1",group.by = "MULTI_Sample")
FeaturePlot(sbmacs,"C1qa",reduction = "umap1")
```

```{r}
sbmacs <- SetIdent(sbmacs, value = sbmacs[["MULTI_Sample"]])
DimPlot(sbmacs,reduction = "umap1")
sbmacs <- FindNeighbors(sbmacs, dims = 1:50)

```

```{r}
sbmacs <- FindClusters(sbmacs, resolution = 0.3,method = 1)
DimPlot(sbmacs,reduction = "umap1")
DimPlot(sbmacs,reduction = "umap1",group.by = "MULTI_Sample")
VlnPlot(sbmacs,"H2-Ab1")
FeaturePlot(sbmacs,reduction = "umap1",features = "H2-Ab1")
VlnPlot(sbmacs,"H2-Ab1",split.by = "MULTI_Sample")
```


### Removing manually outlier cells (clustered as macrophages but beeing distant from them)
```{r}

plot <- DimPlot(object = sbmacs,reduction = "umap1")
# Follow instructions in the terminal to select points
cells.located <- CellSelector(plot = plot)
cells.located
# Automatically set the identity class of selected cells and return a new Seurat object
sbmacs2 <- subset(sbmacs)
sbmacs2 <- CellSelector(plot = plot, object = sbmacs2, ident = 'SelectedCells')


sbmacs3 <- subset(sbmacs2,idents = "SelectedCells")
DimPlot(sbmacs3,reduction = "umap1")
```


### Subsetting to have MM 2- PBS with MM 2- & 2+ LPS
```{r}
sbmacs4 <- SetIdent(sbmacs3,value = sbmacs3[["RNA_snn_res.0.3"]])
sb2neg <- subset(sbmacs4, ident = c("0","2","3"))
DimPlot(sb2neg,reduction = "umap1")
```
```{r}
sb2neg <- ScaleData(sb2neg,do.center = F)
sb2neg <- NormalizeData(sb2neg)
sb2neg <- FindVariableFeatures(sb2neg,selection.method = "mvp")
sb2neg <- RunPCA(sb2neg)
ElbowPlot(sb2neg,50)
for (i in 2:50){
  sb2neg2 <- RunUMAP(sb2neg,dims = 1:i)
  print(DimPlot(sb2neg2))}
```
```{r}
sb2neg2 <- RunUMAP(sb2neg,dims = 1:14,reduction.name = "umap.tr")
DimPlot(sb2neg2,reduction = "umap.tr")
```

### Subsetting to have MM 2+ PBS with MM 2- & 2+ LPS
```{r}
sb2pos <- subset(sbmacs4, ident = c("1","2","3"))
DimPlot(sb2pos,reduction = "umap1")
```
```{r}
sb2pos <- ScaleData(sb2pos,do.center = F)
sb2pos <- NormalizeData(sb2pos)
sb2pos <- FindVariableFeatures(sb2pos,selection.method = "mvp")
sb2pos <- RunPCA(sb2pos)
ElbowPlot(sb2pos,50)
for (i in 2:50){
  sb2pos2 <- RunUMAP(sb2pos,dims = 1:i)
  print(DimPlot(sb2pos2))}
```

```{r}
sb2pos2 <- RunUMAP(sb2pos,dims = 1:20,reduction.name = "umap.tr")
DimPlot(sb2pos2,reduction = "umap.tr")
```


## Monocle3 analyses

### NEG
```{r}
genes <- rownames(sb2neg2@assays$RNA@data)
gene_annotation <- as.data.frame(genes)
```

```{r}
rownames(gene_annotation) <- genes
colnames(gene_annotation) <- 'gene_short_name'
```


```{r}
cds <- new_cell_data_set( sb2neg2@assays$RNA@data,
cell_metadata = sb2neg2@meta.data, gene_metadata = gene_annotation)
```


```{r}
partition <- c(rep(1, length(cds@colData@rownames)))
names(partition) <- cds@colData@rownames
partition <- as.factor(partition)
cds@clusters@listData[["UMAP"]][["partitions"]] <- partition
```

```{r}
seurat_clusters <- sb2neg2@meta.data$RNA_snn_res.0.3
names(seurat_clusters) <- sb2neg2@assays[["RNA"]]@data@Dimnames[[2]]
cds@clusters@listData[["UMAP"]][["clusters"]] <- seurat_clusters
cds@int_colData@listData$reducedDims@listData[["UMAP"]] <- sb2neg2@reductions[["umap.tr"]]@cell.embeddings
```

```{r}
cds <- learn_graph(cds, use_partition = F, learn_graph_control = list(minimal_branch_len = 20))
plot_cells(cds, color_cells_by = 'cluster', label_groups_by_cluster=TRUE, label_leaves=FALSE, label_branch_points=FALSE, graph_label_size=0,cell_size = 1.5,group_label_size = 9,label_cell_groups = my_color_palette[c(1,3,4)],)
```

```{r}
cds <- order_cells(cds, reduction_method = "UMAP")
plot_cells(cds, x=1, y=2, reduction_method = "UMAP", cell_size = 0.95, color_cells_by = "pseudotime", label_cell_groups=F,
label_leaves=FALSE, label_branch_points=FALSE, graph_label_size=4, trajectory_graph_color = "green", trajectory_graph_segment_size = 1.05, label_roots = FALSE,group_label_size = 9,)
```


### POS
```{r}
genes <- rownames(sb2pos2@assays$RNA@data)
gene_annotation <- as.data.frame(genes)
```

```{r}
rownames(gene_annotation) <- genes
colnames(gene_annotation) <- 'gene_short_name'
```


```{r}
cds <- new_cell_data_set( sb2pos2@assays$RNA@data,
cell_metadata = sb2pos2@meta.data, gene_metadata = gene_annotation)
```


```{r}
partition <- c(rep(1, length(cds@colData@rownames)))
names(partition) <- cds@colData@rownames
partition <- as.factor(partition)
cds@clusters@listData[["UMAP"]][["partitions"]] <- partition
```

```{r}
seurat_clusters <- sb2pos2@meta.data$RNA_snn_res.0.3
names(seurat_clusters) <- sb2pos2@assays[["RNA"]]@data@Dimnames[[2]]
cds@clusters@listData[["UMAP"]][["clusters"]] <- seurat_clusters
cds@int_colData@listData$reducedDims@listData[["UMAP"]] <- sb2pos2@reductions[["umap.tr"]]@cell.embeddings
```

```{r}
cds <- learn_graph(cds, use_partition = F, learn_graph_control = list(minimal_branch_len = 20))
plot_cells(cds, color_cells_by = 'cluster', label_groups_by_cluster=TRUE, label_leaves=FALSE, label_branch_points=FALSE, graph_label_size=4,cell_size = 1.5,group_label_size = 9)
```

```{r}
cds <- order_cells(cds, reduction_method = "UMAP")
plot_cells(cds, x=1, y=2, reduction_method = "UMAP", cell_size = 0.95, color_cells_by = "pseudotime", label_cell_groups=T,
label_leaves=FALSE, label_branch_points=FALSE, graph_label_size=4, trajectory_graph_color = "green", trajectory_graph_segment_size = 1.05, label_roots = FALSE,group_label_size = 9)
```
